
** title: "Practical ML" **
** author: "Sarani" **
** date: "March 7, 2019" **
** output: html_document **

** Coursera Final Assignment **


** Library and Data Access **

```{r}

library(knitr)
library(caret)
library(rpart)
#install.packages("rpart.plot")
library(rpart.plot)
#install.packages("rattle")
library(rattle)
#install.packages("randomForest")
library(randomForest)
#install.packages("e1071")
library(e1071)
set.seed(234)


Trainlink <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
Testlink  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"



df_training <- read.csv(url(Trainlink))
df_testing  <- read.csv(url(Testlink))
```

** Data Treatment **

```{r}
# create a partition with the training dataset 
inTrain  <- createDataPartition(df_training$classe, p=0.7, list=FALSE)
TrainSet <- df_training[inTrain, ]
TestSet  <- df_training[-inTrain, ]


# Check number of variables and their names
dim(TrainSet)
colnames(TrainSet)

# remove variables with Nearly Zero Variance
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]


# remove variables that have mostly missing values (90%)
MissRemove    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.90
TrainSet <- TrainSet[, MissRemove==FALSE]
TestSet  <- TestSet[, MissRemove==FALSE]

# Check remaining variables
dim(TrainSet)
colnames(TrainSet)

# remove other unnecessary variables
TrainSet <- TrainSet[, -(1:6)]
TestSet  <- TestSet[, -(1:6)]

# Check the final number of variables
dim(TrainSet)
colnames(TrainSet)
```

** Model Fit and Prediction **

```{r}
# model fit
set.seed(234)
RF <- trainControl(method="cv", number=5, verboseIter=FALSE)
fitRF <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=RF)
fitRF$finalModel

# prediction on Test dataset
predictRF <- predict(fitRF, newdata=TestSet)
confRF <- confusionMatrix(predictRF, TestSet$classe)
confRF

# Accuracy is high 0.99 which is good and OOB error is also low 0.64%

# Final prediction
pred_test <- predict(fitRF, newdata=df_testing)
pred_test